% called by main.tex
%
\chapter{Objectives}
\label{ch::chapter2}

\section{General objectives}
The general objectives of this project focus on the understanding and analysis of audio containing infant’s cries and also to advance in the field of infant’s health by providing new knowledge, so that future research and technological developments can improve medical care for newborns. 
\begin{itemize}
    \item To investigate advanced techniques for \textbf{neonatal audio analysis}.
    \item To carry out an exhaustive investigation of the techniques that have been used to date for the classification of crying.
    \item Conduct a comparison of results on Machine Learning models used for neonatal sound classification. 
    \item Collect and verify labelled audios for training models that accurately represent different sounds emitted by infants.
    \item Investigate and apply \textbf{audio processing methods} and features that improve the models' ability to distinguish between infant cries and other sounds. 
    \item Investigate and apply different \textbf{Python libraries} for the different phases of the project. 
    \item Manually tag the videos provided by the HUBU for training the models.
    \item Create diagrams and graphs for a better understanding of the results visually.
\end{itemize}

\section{Technical objectives}
The technical objectives of this project focus on applying the researched methods to the creation of a robust system that can be used in clinical settings.
\begin{itemize}
    \item Implement a robust \textbf{MLP} model for audio analysis in neonatology.
    \item Implement a robust \textbf{SVM} model for audio analysis in neonatology.
    \item Implement a robust \textbf{LSTM} model for audio analysis in neonatology.
    \item Use a version control system with GitHub.
    \item Use a Latex typesetting system for the creation of the report and annexes to guarantee a certain quality in the document.
    \item Use the Overleaf editor to produce the report and annexes. 
\end{itemize}

\newpage
\section{Personal objectives}
The personal objectives of this project reflect my enthusiasm to grow as a researcher through this project and to put into practice the knowledge acquired so far. 
\begin{itemize}
    \item Try to contribute a novel approach or technique to improve the \textbf{detection of diseases in neonates}. 
    \item To put into practice some of the knowledge acquired in my academic career and, above all, the knowledge acquired in this master's degree stage in a real project.
    \item To strengthen my scientific and technical communication skills.
    \item To develop and improve my skills in the analysis and handling of audio data.
    \item Collaborate with professionals in the field of neonatology to validate and improve the developed models.
\end{itemize}